One page description

### 2016-10-03 publications/Journal/1-introduction.tex
Small changes to section 1

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
Whereas \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, its inherent parallelism is often insufficient~\cite{Himmelspach}.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to \textsf{DEVS} simulation~\cite{globaltimewarp}.
Through synchronization protocols, different simulation cores can be at different points in simulated time, increasing parallelism significantly, at the cost of synchronization overhead.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other in specific models~\cite{Jafer}.
The applicability of parallel simulation capabilities of current tools is therefore constrained to specific domains.

This paper introduces \textit{DEVS-Ex-Machina}\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``\textit{dxex}''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model: users only determine which protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the same model with different synchronization protocols.
Since the ideal synchronization protocol might change throughout the simulation, we also support runtime switching between these protocols.
This runtime switching can be based on performance metrics, which are logged during simulation.
Information is made available to a separate component, where a choice can be made about which synchronization protocol to use next. % It's manual now, can be extended to full automatic.
Additionally, we investigate in this paper how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component to our simulation kernel.

Our tool is based on \textit{PythonPDEVS}, but implemented in C++11 for increased performance, using features from the new C++14 standard when supported by the compiler.
Unlike \textit{PythonPDEVS}, \textit{dxex} only supports multicore parallelism, thus no distributed simulation.

We implemented a model that, depending on a single parameter, changes its ideal synchronization protocol.
Using several models, we demonstrate the factors influencing the performance under a given synchronization protocol.
\textit{Dxex}, then, is used to compare simulation using exactly the same tool, but with a varying synchronization protocol.
With \textit{dxex} users can always opt to use the fastest protocol available, and through its modularity, users could even implement their own.
To verify that this modularity does not hinder performance, we compare to another \textsf{Parallel DEVS} simulation kernel, called \textit{adevs}~\cite{adevs}.

%Usage of Hyperref with hardcoded label is hardly ideal, but ~\ref doesn't work if the layout disables numbered sections, so a hack is needed. The alternative is \nameref but that isn't all that readable.
The remainder of this paper is organized as follows:
Section~\textsc{\nameref{sec:2-background}} introduces the necessary background on synchronization protocols.
Section~\textsc{\nameref{sec:3-features}} elaborates on our design that enables this flexibility.
In Section~\textsc{\nameref{sec:4-performance}}, we evaluate performance of our tool by comparing its different synchronization protocols, and compare to \textit{adevs}.
We continue by introducing runtime switching of synchronization protocols and different options for model allocation in Section~\textsc{\nameref{sec:4b-hotswap}} and Section~\textsc{\nameref{sec:4a-allocation}}, respectively.
Related work is discussed in Section~\textsc{\nameref{sec:5-related-work}}.
Section~\textsc{\nameref{sec:6-conclusion}} concludes the paper and gives future work.

>>>>

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
Whereas \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, its inherent parallelism is often insufficient~\cite{Himmelspach}.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to (\textsf{Parallel}) \textsf{DEVS} simulation~\cite{globaltimewarp}.
With synchronization protocols, different simulation cores can be at different points in simulated time, significantly increasing parallelism, at the cost of synchronization overhead.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other for specific models~\cite{Jafer}.
The parallel simulation capabilities of current tools are therefore limited to specific domains.

This paper introduces \textit{DEVS-Ex-Machina}\footnote{\url{https://bitbucket.org/bcardoen/devs-ex-machina}} (``\textit{dxex}''): our simulation tool~\cite{dxex} which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model: users only determine the protocol to use.
Users who simulate a wide variety of models, with different ideal synchronization protocols, can run the exact same model using the exact same tool, but with different synchronization protocols.
As model behaviour, and thus the ideal synchronization protocol, might change during simulation, runtime switching of synchronization protocols is also supported.
This runtime switching can be based on performance metrics, which are logged during simulation.
Information is made available to a separate component, where a choice can be made about which synchronization protocol to use. % It's manual now, can be extended to full automatic.
Additionally, we investigate how model allocation influences the performance of our synchronization protocols.
To this end, we have included an allocation component in our simulation kernel.

Our tool is based on \textit{PythonPDEVS}~\cite{PythonPDEVS}, but implemented in C++11 for increased performance, using features from the new C++14 standard when supported by the compiler.
Unlike \textit{PythonPDEVS}, \textit{dxex} only supports multicore parallelism, thus no distributed simulation.

Using several benchmark models, we demonstrate the factors influencing the performance under a given synchronization protocol.
Additionally, we investigate a model which changes its behaviour (and ideal synchronization protocol) during simulation.
\textit{Dxex}, then, is used to compare simulation using exactly the same tool, but with a varying synchronization protocol.
With \textit{dxex} users can always opt to use the fastest protocol available, and through its modularity, users could even implement their own, or variants of existing ones.
To verify that this modularity does not hamper performance, we compare to \textit{adevs}~\cite{adevs}, another \textsf{Parallel DEVS} simulation tool.

The remainder of this paper is organized as follows:
Section~\textsc{\nameref{sec:2-background}} introduces the necessary background on synchronization protocols.
Section~\textsc{\nameref{sec:3-features}} elaborates on our design that enables this flexibility.
In Section~\textsc{\nameref{sec:4-performance}}, we evaluate performance of our tool by comparing its different synchronization protocols, and compare to \textit{adevs}.
We continue by introducing runtime switching of synchronization protocols and different options for model allocation in Section~\textsc{\nameref{sec:4b-hotswap}} and Section~\textsc{\nameref{sec:4a-allocation}}, respectively.
Related work is discussed in Section~\textsc{\nameref{sec:5-related-work}}.
Section~\textsc{\nameref{sec:6-conclusion}} concludes the paper and gives future work.
